---
title: "QRM II Graded Assignment (8)"
subtitle: "Period 1, 2026"
author: "Group 23"
date:  "`r format(Sys.Date(), '%d-%m-%Y')`"
geometry: margin=0.8in
output:
  pdf_document:
    number_sections: true
    latex_engine: xelatex
  html_document:
    df_print: paged

header-includes:
- "\\setcounter{section}{-1}"
- |
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{
    showspaces = false,
    showtabs = false,
    breaklines,
    commandchars=\\\{\}
  }
format:
  html:
    code-overflow: wrap
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, fig.width = 6, fig.height = 3, out.width = "55%", fig.align = "center")
```

# Introduction

This assignment is to be completed in groups of 4-5. Further, all students in your group need to be assigned to the same R tutorial group (Friday's tutorial). You can sign yourself up for a group on Canvas. Please do so \textbf{before the start of your first R tutorial on Friday September 4th.} You can use the Discussion Board in Canvas if you do not have a group yet or if your group is incomplete.

The assignment has 5 parts, and each part corresponds to the course material of that week (with the exclusion of week 6, for which there is no R programming material).

You are supposed to hand in these assignments on Canvas at the following dates:

- **Deadline 1** *Sunday September 27th, at 23:59pm*: you are supposed to hand in weeks 1, 2, and 3 of this assignment. This will determine 18% of your overall course grade
- **Deadline 2** *Sunday October 11th, at 23:59pm*: you are supposed to hand in weeks 4, and 5 of this assignment. This will determine 12% of your overall course grade

The R tutorials (each Friday) will consist of two halves. During the first half, you will discuss the tutorial exercises. These can be downloaded separately from Canvas. During the second half, you can work on this graded assignment within your own group. The purpose is that you find out how to work with R for doing statistical analyses by yourself. The tutorial exercises are meant to teach you basic commands to get you started, but to answer the problem sets in this assignment, you might need to research your own solutions, and use functions and commands not described in the tutorial exercises. Learning how to solve your own research problems is integral part of learning R. When you and your group get stuck on how to approach an exercise, the hierarchy in finding your way is as follows:

- use the concepts from the tutorial exercises;
- use the cheat sheets available on Canvas;
- use Google, YouTube, StackOverflow, or another website;
- ask the teacher.

The use of generative AI is \textbf{not} permitted and may result in a grade of 0. See the AI protocol in the course manual for details.

To answer the assignment, you can simply fill out this R markdown document. There are designated places which you can fill with R code. There are also designated spaces for you to answer each question. Often, the structure of an answer will be as follows. First, you type the R code in the designated box. This will show how you analyzed the data to get the answer to the question. Below the box for the R code, you will then summarize your answer to the question, i.e. what are the conclusions that you draw from the data analysis?

When handing in, you are supposed to submit this .Rmd file, and a knitted version of this document. You can knit this document to pdf, word, or html. Knitting to pdf requires you to have a .tex distribution installed on your computer. Knitting to Word requires you to have Word installed.

The exercises are designed such that you should be able to finish the majority of them during the tutorial each week. If you are not able to finish them fully during that time, you are expected to work on it in your own time using the computers on campus or your own device. It is best to meet as a group in-person when working together. If you want to work remotely, github is a good platform to guarantee smooth collaboration. Alternatively, you can email this .Rmd file back and forth to one another as a group, but this is not recommended as it is more cumbersome.

We encourage you to keep your code blocks, printing statements, and final answers, as short as possible. In any case, there is a page limit of 6 pages per week, which encompasses the total length of this document which consists of the questions, your coding lines, and your answers. When your answers to questions of the respective week exceed this page limit, they will not be graded, resulting in zero points.

Each week consists of 1, 2, or 3 subquestions. The total amount of points you can earn per week is 20 points.

# Week 1

1.  Find the dataset “movies3.tsv” on Canvas. Describe your data set: How many observations does it have. How many variables are there? How many subjects? What consists of a subject? \textbf{[4 points]}

```{r}
movies <- read.delim("movies3.tsv")

nrow(movies)                 # Number of rows: 707
ncol(movies)                 # Number of columns: 19
length(unique(movies$title)) # Number of unique titles: 707
```

**Your Answer:**

1.  How many observations? There are 707 observations (707 rows).

2.  How many variables? There are 19 variables

3.  How many subjects? There are 707 subjects, we tested for uniqueness and there were no duplicates.

4.  What consists of a subject? A subject is a single movie. Each row in the dataset represents one movie, described by variables such as: title, budget, revenue, runtime, genre.

<!-- -->

2.  Which of the following types of variables are present in your data set? (i) nominal; (ii) ordinal; (iii); interval; (iv) ratio. If present, name one example of such a variable present in your data set. \textbf{[4 points]}

```{r}
table(movies$genre)
table(movies$original_language)
table(movies$first_actor_gender)
table(movies$director_gender)

class(movies$budget)
summary(movies$budget)
class(movies$runtime)
summary(movies$runtime)
class(movies$release_year)
```

**Your Answer:**

\*i, There are nominal variables, for example the original language.

\*ii: There are no ordinal variables, since none of the categorical variables (e.g. genre, original_language, gender) have a meaningful rank order.

\*iii, There are interval variables, for example the release dates.

\*iv, there are ratio variables for example the runtime.

<!-- -->

3.  A movie studio wants to know which types of movies give maximal profit. Perform the following steps to provide the movie studio with an analysis which corresponds to their request:

<!-- -->

a.  Create the variable profits as the revenue of a movie minus its budget. Report its mean, median, maximum, and minimum. \textbf{[2 points]}
b.  Which movie has the highest profits in your data set and how much are these profits. Which movie has the lowest and how much are its profits? If multiple movies have the exact same highest or lowest profits, give only one example. \textbf{[2 points]}
c.  Create a boxplot of the variable profits. Make sure it has an appropriate title, and appropriate titles and labels for the x- and y-axis. Give Q1, Q2, Q3, and Q4. What does this tell you about the nature of making money in the movies industry? \textbf{[2 points]}
d.  Add a new variable to your data set the log of profits. When creating this variable, what happens to movies for which profits is zero or negative? What then happens when you calculate the mean of log of profits? \textbf{[2 points]}
e.  For movies that have a profit of zero or less, replace log of profits with "NA". What is now the mean of log of profits? Create a boxplot for log of profits, again with an appropriate title, x- and y-axis labels. How does it compare to the boxplot you made under c.)? \textbf{[2 points]}
f.  Create a scatterplot of with the runtime of movies on the x-axis and the average vote of movies on the y-axis. What do you conclude from the scatterplot? Are movies with a longer runtime considered worse or better by the audience, or does the audience not have a preference? Why do you think this is the case? \textbf{[2 points]}

For each step, you should provide first all the code you used to answer the question and then formulate an answer using full sentences.

*Step a*

```{r}

movies$profits <- movies$revenue - movies$budget
mean(movies$profits)
median(movies$profits)
max(movies$profits)
min(movies$profits)

```

**Your Answer:**

The mean profit is \$47,019,054, while the median profit is only \$49,144. The maximum profit is \$1,024,888,979 and the minimum profit is -\$150,000,000. The large gap between the mean and the median indicates that the distribution of profits is strongly right-skewed: most movies earn modest profits, while a small number of very successful movies earn very large profits and pull the mean upwards.

*Step b*

```{r}

movies[which.max(movies$profits), c("title", "profits")]
movies[which.min(movies$profits), c("title", "profits")]

```

**Your Answer:**

The film with the highest profit is *The Lord of the Rings: The Return of the King*, with a profit of \$1,024,888,979. The film with the lowest profit is *The Wolfman*, with a loss of \$150,000,000.

*Step c*

```{r}

boxplot(movies$profits,
        main = "Boxplot of Movie Profits",
        xlab = "Movies",
        ylab = "Profit (USD)")

quantile(movies$profits)

```

**Your Answer:**

The boxplot shows that film profits are widely distributed. Most films generate a profit around the median of \$49,144, which is relatively modest. Notably, the first quartile (Q1) is negative (-\$2,170,357), meaning that at least 25% of films operate at a loss. At the same time, there are many high outliers, reaching a maximum of \$1,024,888,979; this indicates that a small number of films—the "blockbusters"—generate enormous profits, while the rest of the industry earns far less or even incurs losses. This suggests that profitability in the film industry is characterized by high risk and high potential returns: the majority of films are not particularly profitable, but a small number of successful films offset this.

*Step d*

```{r}

# Calculate profits
movies$profits <- movies$revenue - movies$budget

# Calculate the natural logarithm of profits
movies$log_profits <- log(movies$profits)

# Calculate the mean of log profits
mean(movies$log_profits)

```

**Your Answer:**

For films with negative profit, the logarithm cannot be calculated because the logarithm is defined only for positive values; these result in NaN. Consequently, the mean of `log_profits` also becomes NaN, as R cannot calculate a mean while the column contains invalid values.

*Step e*

```{r}

movies$log_profits[movies$profits <= 0] <- NA
mean(movies$log_profits, na.rm = TRUE)

boxplot(movies$log_profits,
        main = "Boxplot of Log of Movie Profits",
        xlab = "Movies",
        ylab = "Log(Profit)")

```

**Your Answer:**

After replacing the log(profits) for films with zero or negative profit with NA, the mean of log(profits) is now 17.395—a valid number, unlike the NaN result obtained in step 3d. Compared to the boxplot of profits from 3c, the boxplot of log(profits) is much more symmetrical: the median lies roughly in the middle of the box, and the extreme upper outliers have disappeared. This is because the log transformation compresses the highly skewed distribution of the original profit figures, making the data better approximate a symmetrical distribution.

*Step f*

```{r}

plot(movies$runtime, movies$vote_average,
     main = "Scatterplot of Runtime vs Average Vote",
     xlab = "Runtime (minutes)",
     ylab = "Average Vote")

```

**Your Answer:**

The scatter plot reveals a number of notable points that are likely data anomalies: a few films with a runtime of 0 minutes (presumably missing data) and some with an average vote of 0 (likely films that have not yet received any votes). Looking at the general pattern in the rest of the data, we see that films with longer runtimes generally receive slightly higher average votes—specifically, films exceeding 150 minutes are clustered at the higher end, whereas shorter films show a wider spread, including lower scores. A possible explanation for this is that longer films have more scope to flesh out the plot and characters, thereby engaging the audience more deeply with the story and leading to higher ratings.

# Week 2

1 Is your dataset movies3.tsv the full population, or is it a sample of a larger population? If the latter, how would you describe the full population? \textbf{[4 points]}

**Your Answer:**

Considering the largest index number exceeds the amount of unique index numbers by more than 4.000, our dataset movies3.tsv is the sample of a larger population. The full popultation we would describe as: "movies".

2

a.  For which actor in your data set do you observe the most movies? \textbf{[2 points]}
b.  What is the average revenue of the movie in which this actor plays and does the revenue lie above or below the revenue of an average movie according to your data set? \textbf{[2 points]}\
c.  How trustworthy do you consider your conclusion to answer 2b? Use the term "law of large numbers" in your explanation. \textbf{[2 points]}

*step a*

```{r}

library(DescTools)

mode_value <- Mode(movies$first_actor, na.rm = TRUE)

print(mode_value)

```

**Your Answer:**

Nicolas Cage

*step b*

```{r}
# Select movies featuring Nicolas Cage
has_cage <- apply(movies, 1, function(row) {
  any(grepl("Nicolas Cage|Nicholas Cage", row, ignore.case = TRUE))
})
cage_movies <- movies[has_cage, ]

# Calculate average revenues
cage_mean <- mean(cage_movies$revenue, na.rm = TRUE)
overall_mean <- mean(movies$revenue, na.rm = TRUE)

cage_mean
overall_mean

comparison <- if (cage_mean > overall_mean) {
  "higher than"
} else if (cage_mean < overall_mean) {
  "lower than"
} else {
  "equal to"
}

```

**Your Answer:**

The average revenue of movies featuring Nicolas Cage is approximately \$119,805,381. This is higher than the average revenue of all movies in the dataset, which is approximately \$77,277,126.

*step c*

```{r}
# Count movies with available revenue data
sum(!is.na(cage_movies$revenue))
sum(!is.na(movies$revenue))

```

**Your Answer:**

The conclusion in 2b describes this dataset, but it may not reliably apply to all movies. Nicolas Cage’s average revenue is based on only 7 movies, compared with 707 movies for the overall average. A few very successful movies could strongly affect his average. According to the law of large numbers, the sample mean tends to approach the population mean as the number of independent, representative observations increases. A larger, representative sample of his movies would therefore make the conclusion more reliable.

3 For this question, you will assume that your data set is the full population.

a.  Recode profits such that it is expressed in millions. What is the variance of the variable profits (in millions) in your data set? \textbf{[2 points]}
b.  Create a new data set, called movies_sample. Make sure that it is a random sample of your data set of 25 movies. What is the variance of profits in this random sample? How does it compare to the variance of profits in 2a? \textbf{[2 points]}
c.  In a for loop, create 100 different samples of 25 movies, as in b, and estimate the variance within each sample. Save the variance of each sample in a vector called sample_vars. So the first position of the vector would have the variance of the first sample, the second position the variance of the second sample, etc. Print the start of this vector. \textbf{[2 points]}
d.  Summarize and make a histogram of sample_vars. What is the mean, standard deviation and shape of its distribution? \textbf{[2 points]}
e.  In your opinion, is a sample of 25 movies sufficient to get a reliable estimate of the population variance of profits, using the sample variance? Explain? \textbf{[2 points]}

*step a*

```{r}
movies$profits <- movies$revenue - movies$budget
movies$profits_million <- movies$profits / 1000000

population_var <- mean(
  (movies$profits_million - mean(movies$profits_million))^2
)

population_var
```

**Your Answer:**

Profits were converted to millions by dividing by 1,000,000. The population variance is approximately 15,057.39 (million dollars)². Since the dataset represents the full population, the variance was calculated using N rather than N − 1.

*step b*

```{r}
set.seed(123)

movies_sample <- movies[
  sample(nrow(movies), 25, replace = FALSE),
]

sample_var <- var(movies_sample$profits_million)
sample_var

```

**Your Answer:**

A random sample of 25 movies was selected. The sample variance of profits is 39,218.46 (million dollars)², which is approximately 2.6 times the population variance of 15,057.39. This sample therefore overestimates the population variance due to random sampling variation.

*step c*

```{r}
set.seed(456)

sample_vars <- numeric(100)

for (i in 1:100) {
  sample_movies <- movies[
    sample(nrow(movies), 25, replace = FALSE),
  ]

  sample_vars[i] <- var(sample_movies$profits_million)
}

head(sample_vars)

```

**Your Answer:**

A for-loop was used to draw 100 random samples of 25 movies. The variance of profits in millions was calculated for each sample and stored in the vector sample_vars. The first six variances are 2036.467, 3982.323, 51134.598, 14837.488, 14872.725, and 9271.950.

*step d*

```{r}
summary(sample_vars)
mean(sample_vars)
sd(sample_vars)

hist(
  sample_vars,
  main = "Distribution of sample variances",
  xlab = "Sample variance (million dollars squared)",
  col = "lightblue",
  breaks = 15
)

```

**Your Answer:**

The mean of the sample variances is 15,686.51, and the standard deviation is 14,582.10. The distribution is strongly right-skewed: most sample variances are relatively low, while a few large values create a long right tail.

*step e*

**Your Answer:**

A sample of 25 movies does not appear sufficient for a reliable estimate of the population variance. The sample variances vary widely, with a standard deviation of 14,582.10. Although their mean is close to the population variance of 15,057.39, an individual sample can give a very different estimate. A larger sample would generally provide a more precise estimate.

# Week 3

For the next part of the assignment, assume that the movies in your data frame are a random sample of a larger population of movies.

1

a.  Create a new data set that only includes movies that are of the genre "Thriller". For these thriller movies, give a 99 percent confidence interval for the variable *runtime*. Interpret the result. \textbf{[2 points]}
b.  Now, assume that the variance of *runtime* amongst thriller movies in your data is exactly the same as the variance of *runtime* in the population. Under this assumption, give a 99 percent confidence interval for the variable *runtime* among thriller movies. Interpret the result. Is you confidence interval wider or less wide than the one you found under question 1a? Why is that the case? \textbf{[2 points]}

*step a*

```{r}
# Select thriller movies
thriller_movies <- subset(movies, genre == "Thriller")

# Calculate a 99% confidence interval for the mean runtime
runtime_test <- t.test(thriller_movies$runtime, conf.level = 0.99)

# Show the mean and confidence interval
runtime_test$estimate
runtime_test$conf.int


```

**Your Answer:**

The sample mean runtime of thriller movies is 104.18 minutes. The 99% confidence interval for the population mean runtime is [99.91, 108.44] minutes. We are 99% confident that this interval contains the true mean runtime of thriller movies, assuming the sample is independent and representative. We use a t-distribution because the population standard deviation is unknown.

*step b*

```{r}

n <- nrow(thriller_movies)
xbar <- mean(thriller_movies$runtime)
sigma <- sd(thriller_movies$runtime)

z <- qnorm(0.995)

lower <- xbar - z * sigma / sqrt(n)
upper <- xbar + z * sigma / sqrt(n)

c(lower, upper)


```

**Your Answer:**

The 99% confidence interval = [100.00, 108.35] This means that we are 99% sure that the true mean runtime of thriller movies in the population is between 100.00 and 108.35 minutes.

This interval is slightly narrower than the interval in 1a. In 1a the population variance is unknown, so the t-distribution is used. In 1b the population variance is assumed to be known, so we use the normal distribution. This gives a narrower confidence interval because the values are smaller.

2

a.  Using an appropriate five-step procedure, set up a test for the null hypothesis that the variance of runtime equals $500$. Clearly state your null hypothesis, alternative hypothesis your test statistic, your critical value, and your conclusion. \textbf{[2 points]}
b.  For the validity of your test in 2a, what assumption about the distribution of runtime needs to hold? Make an appropriate plot to test this assumption. What do you conclude? \textbf{[2 points]}

*step a*

```{r}

runtime <- na.omit(movies$runtime)

n <- length(runtime)
s2 <- var(runtime)
df <- n - 1

test_stat <- df * s2 / 500

lower <- qchisq(0.025, df)
upper <- qchisq(0.975, df)

test_stat
lower
upper


```

**Your Answer:**

The null hypothesis is that the population variance of runtime is 500, so (H_0:  sigma\^2 = 500). The alternative hypothesis is that the variance is different from 500, so (H_1:  sigma\^2  neq 500).

For this we used the chi-square test with a 5% significance level. The test statistic is 651.85, the lower critical value is 634.26 and the upper critical value is 781.52.

Because the test statistic falls between these two critical values, the null hypothesis could be true. This means that there is not enough evidence to say that the population variance of runtime is different from 500.

*step b*

```{r}
runtime <- movies$runtime[!is.na(movies$runtime)]

mu <- mean(runtime)
sigma <- sd(runtime)

hist(runtime, breaks = 30, prob = TRUE,
     main = "Histogram of Runtime with Normal Curve",
     xlab = "Runtime")

curve(dnorm(x, mean = mu, sd = sigma),
      col = "red", lwd = 2, add = TRUE)


```

**Your Answer:**

Runtime should be normally distributed. We used a histogram and added a normal curve to check this. The distribution does not follow the normal curve good, especially in the tails. So we can conclude that the normality assumption does not seem to hold very well.

3.  There is an argument going on in the movie studio. *Bob* claims that they should make higher-quality movies, as this will bring in more profits. *Chantal* disagrees. She tells Bob that mediocre movies bring in the most profits. You are asked to advise on who is right.

<!-- -->

a.  Create a new variable called vote_average_rounded. Make sure this variable is the same as vote_average, but without any decimals (i.e., a 6.3 becomes a 6, a 8.7 an 8, etc.). Display a histogram of vote_average_rounded. \textbf{[2 points]}
b.  Create a scatter plot with vote_average_rounded on the x axis and the mean of profits within each category of vote_average_rounded on the y-axis. Make sure it has an appropriate title, and appropriate titles and labels for the x- and y-axis. At which rating of movies are profits the highest? \textbf{[3 points]}
c.  Recreate the scatter plot with year on the x axis and mean_profits on the y-axis, but now add bars around each point, indicating the 95% confidence interval. \textbf{[3 points]}
d.  Write an advice to settle the argument between Bob and Chantal. \textbf{[4 points]}

*step a*

```{r}
# Remove decimals
movies$vote_average_rounded <- floor(movies$vote_average)

# Calculate profits
movies$profits <- movies$revenue - movies$budget

# Plot the distribution of ratings
hist(
  movies$vote_average_rounded,
  breaks = seq(-0.5, 10.5, by = 1),
  main = "Distribution of Movie Ratings",
  xlab = "Rating (decimals removed)",
  ylab = "Number of movies",
  col = "lightblue"
)


```

**Your Answer:**

I created vote_average_rounded by removing the decimal part of vote_average using floor(). For example, 6.3 becomes 6 and 8.7 becomes 8. The histogram shows that most movies have ratings of 5 or 6, while very low and very high ratings are less common. *step b*

```{r}
movies$profits <- movies$revenue - movies$budget

mean_profits <- aggregate(profits ~ vote_average_rounded,
                          data = movies,
                          FUN = mean,
                          na.rm = TRUE)

mean_profits

plot(mean_profits$vote_average_rounded,
     mean_profits$profits / 1000000,
     main = "Mean Profit by Movie Rating",
     xlab = "Rounded Average Vote",
     ylab = "Mean Profit (millions)",
     pch = 19)

mean_profits[which.max(mean_profits$profits), ]


```

**Your Answer:**

Movies with a vote_average_rounded of 8 have the highest average profit: approximately \$313.68 million. The scatter plot shows that higher-rated movies generally have higher average profits in this dataset.

*step c*

```{r}
# Create summary statistics for each rounded rating
profit_summary <- aggregate(
  profits ~ vote_average_rounded,
  data = movies,
  FUN = function(x) c(
    mean = mean(x, na.rm = TRUE),
    sd = sd(x, na.rm = TRUE),
    n = sum(!is.na(x))
  )
)

# Convert the results into separate columns
profit_summary <- data.frame(
  vote_average_rounded = profit_summary$vote_average_rounded,
  mean_profits = profit_summary$profits[, "mean"],
  sd = profit_summary$profits[, "sd"],
  n = profit_summary$profits[, "n"]
)

# Calculate standard error
profit_summary$se <- profit_summary$sd / sqrt(profit_summary$n)

# Calculate critical t-value for a 95% confidence interval
profit_summary$t_critical <- qt(
  0.975,
  df = profit_summary$n - 1
)

# Calculate lower and upper bounds of the confidence interval
profit_summary$lower_ci <- profit_summary$mean_profits -
  profit_summary$t_critical * profit_summary$se

profit_summary$upper_ci <- profit_summary$mean_profits +
  profit_summary$t_critical * profit_summary$se

# Plot mean profits in millions
plot(profit_summary$vote_average_rounded,
     profit_summary$mean_profits / 1000000,
     main = "Mean Profit by Movie Rating with 95% Confidence Intervals",
     xlab = "Rounded Average Vote",
     ylab = "Mean Profit (millions)",
     pch = 19,
     ylim = range(c(profit_summary$lower_ci,
                    profit_summary$upper_ci) / 1000000,
                  na.rm = TRUE))

# Add 95% confidence interval bars
arrows(x0 = profit_summary$vote_average_rounded,
       y0 = profit_summary$lower_ci / 1000000,
       x1 = profit_summary$vote_average_rounded,
       y1 = profit_summary$upper_ci / 1000000,
       angle = 90,
       code = 3,
       length = 0.05)

 


```

**Your Answer:**

This Scatterplot gives a more clear view on how the profits are distributed within movies with a different rounded average vote. One thing that is really noticeable is the really large 95% interval of rounded average vote of 8 compared to the other rounded average vote. This is mostly due to the low frequency of movies with a vote of 8, which is shown in step a.

*step d*

The results provide mostly support for Bob because movies with the highest average vote has the highest mean profit. However the movies with a rounded average of 8 have a extremely wider confidence interval. This means that its estimate is really uncertain. Movies with more moderate ratings have a narrower confidence interval, so their mean profits can be estimated more precisely, but their mean profit is lower. Therefore, we cannot fully conclude whether Bob or Chantal is fully right, although the results in general are more consistent with Bob's arguments.

# Week 4

1.  There is another argument going on in the movie studio. *Bob* claims that production budgets are getting out of hand, and that the studio should focus on making cheaper movies. *Chantal* disagrees. She tells Bob that \`\`Every dollar we spend on movie production is more than offset by the increase in movie profits''.

<!-- -->

a.  Set up a regression model to test Chantal's claim, and estimate it. That is, estimate: $$\text{Profits}_i=\beta_0+\beta_1 \text{Budget}_i +\varepsilon_i.$$ Print a summary of your estimated model. \textbf{[2 points]}
b.  What is the estimated value of $\beta_1$ and how do you interpet it? \textbf{[2 points]}
c.  Test for the null hypothesis that $\beta_1 \geq 0$. Report the p-value and state your conclusion. \textbf{[2 points]}
d.  Next, estimate the model $$\text{Log Profits}_i=\beta_0+\beta_1 \text{Log Budget}_i +\varepsilon_i.$$ When creating the variables Log Profits and Log Budget, make sure that movies with a Revenue or Budget of zero are assigned the value "NA". Print a summary of your estimated model \textbf{[2 points]}
e.  What is the estimated value of $\beta_1$ and how do you interpet it? \textbf{[2 points]}
f.  Which model has better fit? The level-level model or the log-log model? Explain. \textbf{[2 points]}
g.  Who do you think is correct? Bob or Chantal? What would you advise the movie studio to do? \textbf{[2 points]}

*step a*

```{r}
movies$profits <- movies$revenue - movies$budget

model_level <- lm(profits ~ budget, data = movies)

summary(model_level)

```

**Your Answer:**

This model estimates how much profit changes on average when the budget increases by one dollar.

*step b*

```{r}
coef(model_level)["budget"]
```

**Your Answer:**

The estimated value of $\beta_1$ is 1.576. This means that an one dollar increase in production budget is associated with an 1.576 dollar increase in estimated profit.

*step c*

```{r}
t_value <- summary(model_level)$coefficients["budget", "t value"]
t_value
pt(t_value, df = model_level$df.residual)
```

**Your Answer:** The estimated t-statistic is 16.825 and the one-sided p-value for the test $H_0:\beta_1\geq0$ against $H_1:\beta_1<0$ is approximately 1.00. So, we do not reject the null hypothesis.

*step d*

```{r}
# Create log variables
movies$log_profits <- ifelse(
  movies$profits <= 0 | movies$revenue == 0 | movies$budget == 0,
  NA,
  log(movies$profits)
)

movies$log_budget <- ifelse(
  movies$revenue == 0 | movies$budget == 0,
  NA,
  log(movies$budget)
)

# Estimate log-log regression
model_log <- lm(log_profits ~ log_budget, data = movies)

summary(model_log)

```

**Your Answer:**

Movies for which profits are zero or negative are assigned NA because their logarithm cannot be used in the regression. The model is estimated using the remaining observations

*step e*

```{r}
coef(model_log)["log_budget"]
```

**Your Answer:**

The estimated value of $\beta_1$ is 0.48224 since this is log-log model, the coefficient is interpreted as an elasticity. This means that a 1% increase means a 0.482% increase in profits on average.

*step f*

```{r}
## R-squared of level-level model
summary(model_level)$r.squared
summary(model_level)$adj.r.squared

# R-squared of log-log model
summary(model_log)$r.squared
summary(model_log)$adj.r.squared

```

**Your Answer:**

The level-level model has a higher R^2 of 0.2865 than the 0.2353 of the log-log model. Also the adjusted R^2 of level-level model of 0.2855 is higher than the 0.2330 of the log-log model. However the models have different dependent variables so it doesn't give the most accurate statistics. 

*step g*

**Your Answer:**

The results give more support for Chantal than for Bob. In the level-level model, one extra dollar is associated with 1.576 higher profit on average. The coefficient is positive and that means there is no evidence from the one-sided test that higher budgets reduce profits. The log-log model also shows a positive relationship between budget and profits because a 1% increase in budget is associated with an 0.482% increase in profits. Therefore, the results do not support that the studio should focus on making cheap movies. I would advise the studio to not reduce budgets. 

2

a.  Make a plot with a 95% confidence interval with the mean log of budget on the y-axis, and whether the first actor of the movie is male or female on the x-axis. What do you conclude? \textbf{[2 points]}
b.  Estimate the following simple OLS model: $log(budget)_i=\beta_0+\beta_1 \text(FirstActorMale)_i + \varepsilon_i.$ Is the estimated coefficient for $\beta_1$ significantly different from zero? How do you interpret its estimate, and how does this relate to your conclusion in 2a? \textbf{[2 points]}
c.  Now, have a close look at your data frame. Can you find any instances of male first actors who are wrongly labeled as being female, or vice versa? What would such mislabelling mean for the coefficient you estimated under 2b? \textbf{[2 points]}

*step a*

```{r}
movies$log_budget2 <- ifelse(movies$budget > 0, log(movies$budget), NA)

gender_summary <- aggregate(log_budget2 ~ first_actor_gender, data = movies,
  FUN = function(x) c(mean = mean(x), sd = sd(x), n = length(x)))
gender_summary <- data.frame(
  gender = gender_summary$first_actor_gender,
  mean   = gender_summary$log_budget2[, "mean"],
  sd     = gender_summary$log_budget2[, "sd"],
  n      = gender_summary$log_budget2[, "n"])
gender_summary$se    <- gender_summary$sd / sqrt(gender_summary$n)
gender_summary$lower <- gender_summary$mean - qt(0.975, gender_summary$n - 1) * gender_summary$se
gender_summary$upper <- gender_summary$mean + qt(0.975, gender_summary$n - 1) * gender_summary$se
gender_summary

plot(1:2, gender_summary$mean, xlim = c(0.5, 2.5),
     ylim = range(gender_summary$lower, gender_summary$upper),
     xaxt = "n", pch = 19,
     main = "Mean log(budget) by gender of first actor (95% CI)",
     xlab = "Gender of first actor", ylab = "Mean log(budget)")
axis(1, at = 1:2, labels = gender_summary$gender)
arrows(1:2, gender_summary$lower, 1:2, gender_summary$upper,
       angle = 90, code = 3, length = 0.05)
```

**Your Answer:**

The group with a male first actor has the highest mean log(budget): 16.73, compared with 16.35 for a female first actor. This difference of about 0.38 on the log scale means that films with a male first actor have, on average, a budget that is about exp(0.38) $\approx$ 1.46 times as large, i.e. roughly 46% higher.

The two 95% confidence intervals partly overlap. The interval for the female group is also wider, because that group is smaller (n = 143 versus n = 397), which gives a larger standard error.

Conclusion: in this sample, films with a male first actor have a higher mean budget, but because the intervals overlap, the plot alone does not clearly establish that the difference exists in the population. It could also be due to chance. Overlapping intervals are not a formal test, so whether the difference is significant is tested formally in 2b. Note also that this is an association, not a causal effect: factors such as genre, release year and studio may influence both the budget and the gender of the first actor.

*step b*

```{r}
movies$first_actor_male <- ifelse(movies$first_actor_gender == "male", 1, 0)
model_gender <- lm(log_budget2 ~ first_actor_male, data = movies)
summary(model_gender)
confint(model_gender)
```

**Your Answer:**

The regression yields log_budget2 = 16.350 + 0.381 × first_actor_male. $\beta_1 = 0.381$ is exactly the difference between the two group means in 2a (16.73 − 16.35). On average, films with a male lead actor have a budget that is approximately 46% higher (exp(0.381) $\approx$ 1.46).

The difference is not significant at the 5% level (t = 1.879, df = 538, p = 0.061), and the 95% confidence interval [−0.02; 0.78] includes 0; therefore, we do not reject $H_0: \beta_1 = 0$.

Conclusion: In this sample, films with a male lead actor have a higher average log(budget), but the difference is not significant, so we cannot conclude that it exists in the population. This is consistent with the partially overlapping intervals from 2a.

*step c*

```{r}
# Lists used to search for mislabelled actors (run in the Console, they are long):
# sort(unique(movies$first_actor[movies$first_actor_gender == "female"]))
# sort(unique(movies$first_actor[movies$first_actor_gender == "male"]))

# Examples of men labelled as female
movies[movies$first_actor %in% c("Morgan Freeman", "Jackie Chan", "Jamie Foxx",
                                 "Taylor Lautner", "Leslie Nielsen", "Cuba Gooding"),
       c("title", "first_actor", "first_actor_gender")]
```

**Your Answer:**

The `first_actor_gender` variable contains incorrect labels: Morgan Freeman, Jackie Chan, Jamie Foxx, Taylor Lautner, Leslie Nielsen, and Cuba Gooding are classified as “female,” even though they are men. Many of these first names are used by both genders (Morgan, Taylor, Jamie, Leslie), suggesting that the gender was likely determined automatically based on the first name.

This affects $\beta_1$. These men are incorrectly included in the “female” group, causing the average log(budget) of that group to be conflated with that of the men. Consequently, the two groups appear more similar, reducing the observed difference. Therefore, $\beta_1 = 0.381$ is likely an underestimate of the true difference, assuming the errors primarily involve men being misclassified as “female.”

# Week 5

a.  Create a plot of the mean profits by month of release. Do you see any indication that month of release matters to the profits of the movie? \textbf{[2 points]}

b.  Estimate an OLS model which has as dependent variable the log of profits of a movie, and as independent variable the log of budget, a dummy for whether the movie was released in english or not, and a linear term for the month of release. Show a summary of the resulting model and interpret each coefficient. \textbf{[4 points]}

c.  Test for the hypothesis that the coefficient that belongs to month of release is zero. \textbf{[2 points]}

d.  Based on your plot in a.) do you consider the choice that month of release enters the model linearly under b.) reasonable? Estimate a specification that allows for a more flexible curve. In this new specification, test for the null hypothesis that month of release does not impact profits. This might require testing multiple terms at once. \textbf{[4 points]}

e.  One executive at the studio wants to time the release of the movie to a specific month of the year such that they can maximize revenue. Based on your model under d.), What would you advise the movie studio regarding the timing of the release of the movie? \textbf{[2 points]}

The movie studio that you work at is releasing a new movie in 2026. It will be an English-spoken Thriller movie with a budget of 40,000,0000.

f.  Estimate a model that is able to predict the revenue of this movie. Give its predicted revenue and include a 99% prediction interval. \textbf{[6 points]}

*step a*

```{r}

mean_profits_month <- aggregate(profits ~ release_month, data = movies, FUN = mean)

barplot(mean_profits_month$profits / 1000000,
        names.arg = mean_profits_month$release_month,
        main = "Mean Profits by Month of Release",
        xlab = "Month of Release",
        ylab = "Mean Profits (millions)")

```

**Your Answer:**

It seems unlikely that the month of release has nothing to do with mean profits, considering the wide disparities across the year.

*step b*

```{r}
movies$english <- ifelse(movies$original_language == "en", 1, 0)

model_b <- lm(
  log_profits ~ log(budget) + english + release_month,
  data = movies,
  subset = is.finite(log_profits) & budget > 0
)

summary(model_b)

```

**Your Answer:**

The intercept is 9.45382: the predicted log profit for a non-English movie with a budget of 1 and release month 0. Since month 0 does not exist, this has little practical meaning.

Holding the other variables constant:

- A 1% higher budget is associated with approximately 0.484% higher profits.
- English-language movies have approximately 20.2% lower profits than non-English movies, calculated as 100 × (exp(-0.22545) − 1). This difference is not statistically significant (p = 0.557).
- A one-month increase in release month is associated with approximately 1.9% higher profits. This relationship is not statistically significant (p = 0.418).

*step c*

```{r}
summary(model_b)$coefficients["release_month", ]

```

**Your Answer:**

We test $H_0: \beta_{\text{month}} = 0$ against $H_1: \beta_{\text{month}} \neq 0$ using a two-sided t-test. The test statistic is 0.811 and the p-value is 0.418.

Since p \> 0.05, we do not reject $H_0$. There is insufficient evidence of a linear relationship between release month and log profits, controlling for log budget and language.

*step d*

```{r}
data_d <- subset(
  movies,
  is.finite(log_profits) & is.finite(budget) & budget > 0 &
    !is.na(english) & !is.na(release_month)
)

# Model with a more flexible (quadratic) curve
model_d <- lm(
  log_profits ~ log(budget) + english +
    release_month + I(release_month^2),
  data = data_d
)

summary(model_d)

# Model without both month terms
model_without_month <- lm(
  log_profits ~ log(budget) + english,
  data = data_d
)

# Joint F-test for both month terms
anova(model_without_month, model_d)

```

**Your Answer:**

A linear relationship does not seem reasonable based on the plot in part A: mean profits show several peaks and dips rather than a steady increase or decrease. We therefore add a squared release-month term to allow for curvature, although a quadratic curve cannot capture every fluctuation.

The estimated model is:

Predicted log profits = 9.151123 + 0.482036 × log(budget) − 0.188268 × english + 0.145982 × release_month − 0.009968 × release_month².

We test $H_0$: both month coefficients are zero against $H_1$: at least one is non-zero.

The joint F-test gives F(2, 330) = 1.2244 and p = 0.2953. Since p \> 0.05, we do not reject $H_0$. There is insufficient evidence that release month is associated with log profits in this quadratic model, controlling for budget and language.

*step e*

```{r}
predictions <- data.frame(
  release_month = 1:12,
  budget = 40000000,
  english = 1
)

predictions$predicted_log_profit <- predict(
  model_d,
  newdata = predictions
)

predictions[
  order(predictions$predicted_log_profit, decreasing = TRUE),
]

```

**Your Answer:**

We interpret the budget of "40,000,0000" in the assignment as \$40,000,000. The model predicts the highest log profits in July, with August a close second. However, the month terms are not jointly statistically significant (p = 0.2953), so there is insufficient evidence to confidently recommend one month over another. The studio should also consider competition, holidays and marketing when choosing a release date. The model predicts log profits rather than revenue directly.

*step f*

```{r}
# Create indicators for English language and thriller genre
movies$english <- ifelse(movies$original_language == "en", 1, 0)
movies$thriller <- as.integer(
  grepl("thriller", movies$genre, ignore.case = TRUE)
)

# Estimate the revenue model
model_f <- lm(
  revenue ~ budget + english + thriller + release_year,
  data = movies
)

summary(model_f)

# Characteristics of the new movie
new_movie <- data.frame(
  budget = 40000000,
  english = 1,
  thriller = 1,
  release_year = 2026
)

# Predicted revenue with a 99% prediction interval
predict(
  model_f,
  newdata = new_movie,
  interval = "prediction",
  level = 0.99
)

```

**Your Answer:**

The predicted revenue for the new film is approximately \$92.0 million. The 99% prediction interval ranges from roughly -\$179.2 million to \$363.1 million, meaning we are 99% confident that the revenue of an individual new film with these characteristics will fall within this range. The prediction is primarily driven by the budget: each additional dollar of budget is associated with approximately \$2.59 in additional revenue (p < 2e-16), whereas the thriller genre is only marginally significant (p = 0.096), and English (p = 0.91) and release_year (p = 0.62) are not significant. It should be noted that the interval is very wide and the lower bound is negative—an impossibility for revenue. This is due to the large residual standard error (approximately \$104 million), which results in high uncertainty regarding individual predictions, and the fact that a linear model does not inherently account for the fact that revenue cannot drop below zero. Furthermore, since we are predicting for 2026 while the data only extends to 2016, we are extrapolating beyond the data range, making the prediction less reliable.
